A comparison of single-cell trajectory inference methods

### 2018-11-12 manuscript/paper.Rmd

---
title: "Accuracy, scalability, robustness and usability of single-cell trajectory inference methods[a][b]"
output: dynbenchmark::pdf_manuscript
---

@@

`www.github.com/dynverse/dynbenchmark/commit/`r git2r::revparse_single(".","HEAD")$sha``
```{r}
n_tools <- nrow(tools)
n_tools_evaluated <- nrow(filter(tools, evaluated))
n_methods_evaluated <- nrow(filter(methods, evaluated, source %in% c("tool", "adaptation", "offtheshelf")))
n_control_methods_evaluated <- nrow(filter(methods, evaluated, source %in% c("control")))
n_datasets_real <- nrow(list_datasets("real"))
n_datasets_synthetic <- nrow(list_datasets("synthetic"))
```
# Abstract

@@

→
# Results
## Trajectory inference methods
<!-- Method characterisation -->
In this study, we performed a comprehensive evaluation of  `r n_methods_evaluated` TI methods (`r ref("fig", "ti_evaluation_overview", "a")`). In order to make the outputs from different methods directly comparable to each other, we developed a common probabilistic model for representing trajectories from all possible sources (`r ref("fig", "ti_evaluation_overview", "b")`). In this model, the overall topology is represented by a network of "milestones", and the cells are placed within the space formed by each set of connected milestones. Although almost every method returned a unique set of outputs, we were able to classify these outputs into 7 distinct groups, and wrote a common output convertor for each of these groups (`r ref("fig", "results_summary", "a")` and `r ref("sfig", "overview_wrapping")`). When required, we also provided prior information to the method, the severity of which varies (`r ref("fig", "results_summary", "a")`). Namely, having to provide a start cell or an end cell is relatively easy to do and will only introduce moderate bias into the model, whereas having to provide a clustering or time course information will likely introduce heavy bias into the model.[c][d]

@@

Methods in which the topology of the model was fixed by the algorithm or by a parameter obtain a much higher stability score. As topology-free methods have more degrees of freedom, the stability of the method is also typically lower ─ with a few exceptions.[e][f]

We will discuss each evaluation criterion in more detail, after which we conclude with guidelines for method users and future perspectives for method developers.
```{r}
add_fig(
  fig_path = result_file("results_summary.pdf", experiment_id = "08-summary"),
  ref_id = "results_summary", 
  caption_main = glue::glue("A characterisation of the {n_methods_evaluated} methods evaluated in this study, and their overall evaluation results."), 
  caption_text = "(a) We characterised the methods according to the wrapper type, their required priors, whether the inferred topology is constrained by the algorithm (fixed) or a parameter (param), and the types of inferable topologies. The methods are grouped vertically based on the most complex trajectory type they can infer. (b) The overall results of the evaluation on four criteria: benchmarking using a reference trajectory on real and synthetic data, scalability with increasing number of cells and features, stability across dataset subsamples, and quality of the implementation.",
  width = 15,
  height = 18
)
```

```{r}
add_fig(
  fig_path = result_file("results_detailed.pdf", experiment_id = "08-summary"),
  ref_id = "results_detailed", 
  caption_main = glue::glue("Detailed results of the four main evaluation criteria: benchmarking, scalability, stability and quality control."), 
  caption_text = glue::glue("(a) The names of the methods. (b) Results of the benchmark across metrics, dataset sources and dataset trajectory types. The performance of a method is generally more stable across dataset sources, but very variable depending on the metric and trajectory type. (c) Predicted execution times for varying numbers of cells and features (# cells × # features). Predictions were made by training a regression model after running each method on bootstrapped datasets with varying numbers of cells and features. (d) Stability results by calculating the average pairwise similarity between models inferred across multiple runs of the same method.  (e) Usability scores of the tool and corresponding manuscript, grouped per category."),
  width = 15,
  height = 18
)
```

```{r}
add_sfig(
  fig_path = result_file("results_suppfig.pdf", experiment_id = "08-summary"),
  ref_id = "results_suppfig", 
  caption_main = glue::glue(""), 
  caption_text = glue::glue(""),
  width = 30,
  height = 18,
  integrate = FALSE
)
```
## Accuracy
<!-- Accuracy methods -->
We defined several possible metrics to compare a prediction to a reference trajectory (`r ref("snote", "metrics")`). Based on an analysis of their robustness and conformity to a set of rules (`r ref("snote", "metrics")`), we chose four metrics each assessing a different aspect of a trajectory (`r ref("fig", "ti_evaluation_overview", "d")`): the topology (Hamming-Ipsen-Mikhailov, `r label_metric("him")`), the quality of clustering of the assignment of cells to branches (`r label_metric("F1_branches")`), the cell positions (`r label_metric("correlation")`) and the quality of differentially expressed features  (`r label_metric("featureimp_wcor")`). The data compendium consisted of both synthetic datasets, which offer the most exact reference trajectory, and real datasets, which provide the highest biological relevance. These real datasets come from a variety of single-cell technologies, organisms, and dynamic processes, and contain several types of trajectory topologies (`r ref("stable", "datasets")`).  Some real datasets were classified as "gold standard" if the reference trajectory was not extracted from the expression data itself, such as via cellular sorting or cell mixing [@tian_scrnaseqmixologybetter_2018]. All other real datasets were classified as "silver standard". For synthetic datasets we used several data simulators, including a simulator of gene regulatory networks using a thermodynamic model of gene regulation[@schaffter_genenetweaversilicobenchmark_2011]. For each simulation, we used a real dataset as a reference, to match its dimensions, number of differentially expressed genes, drop-out rates and other statistical properties[@zappia_splattersimulationsinglecell_2017].

@@

## Scalability
While early TI methods were developed at a time where profiling more than a thousand cells was exceptional, methods now have to cope with hundreds of thousands of cells, and perhaps soon with more than ten million [@svensson_exponentialscalingsinglecell_2018]. Moreover, the recent application of TI methods on multi-omics single-cell data also showcases the increasing demands on the number of features [@cao_jointprofilingchromatin_2018]. To assess the scalability, we ran each method on up- and downscaled versions of five distinct real datasets. We then modelled the running time using a generalized additive model[@wood_thinplateregression_2003] (`r ref("sfig", "scaling", "a")`), and assessed whether the method had sublinear, linear or superlinear time complexity with respect to the number of cells and features (`r ref("sfig", "scaling", "b")`). As a control, we compared the predicted time with the actual time on all benchmarking datasets, and found that these were moderately to highly correlated (0.5-0.9) for almost every method.[g][h]

@@

←
## Implementation quality and user friendliness

While not directly related to the accuracy of the inferred trajectory, it is also important to assess the quality of the implementation and how easy it is for a biological user to use [@_doesyourcode_2018]. We scored each method using a transparent checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and publication into a peer-reviewed journal (`r ref("stable", "qc_scorseheet")`).

We found that most methods fulfilled the basic criteria, such as free availability and basic code quality criteria (`r ref("fig", "results_detailed", "c")` and `r ref("sfig", "qc_overview")`). While recent methods had a slightly better quality score than older methods, several quality aspects were consistently lacking for the majority of the methods (`r ref("sfig", "qc_overview", " right")`) and we believe that these should receive extra attention from developers. Although these outstanding issues cover all five categories, code assurance and documentation in particular are problematic areas, notwithstanding several studies pinpointing these as good practices [@wilson_best_2014; @artaza_top10metrics_2016].

```{r}
add_sfig(
  fig_path = result_file("qc_overview.rds", experiment_id = "03-methods/02-tool_qc"),
  ref_id = "qc_overview", 
  caption_main = "Quality control of trajectory inference methods.", 
  caption_text = glue::glue("Shown is the score given for each method on every item from our quality control score sheet ({ref('stable', 'qc_scoresheet')}). Each aspect of the quality control was part of a category, and each category was weighted so that it contributed equally to the final quality score. Within each category, each aspect also received a weight depending on how often it was mentioned in a set of papers discussing good practices in tool development and evaluation. This is represented in the plot as the height on the y-axis. Top: Average QC score for each method. Right: The average score of each quality control item. Shown into more detail are those items which had an average score lower than 0.5."),
   width = 15,
   height = 18
)
```

```{r}
add_stable(
 result_file("qc_scoresheet.xlsx", "03-methods/02-tool_qc"),
  "qc_scoresheet",
paste0("→Scoring sheet for method quality control. Each quality aspect was given a weight based on how many times it was mentioned in a set of articles discussing best practices for tool development.←"),
  ""
)
```

@@

________________
# Online Methods
## Data and code availability
→ 
The processed real and synthetic datasets used in this study are deposited on Zenodo ([doi.org/10.5281/zenodo.1211533](https://doi.org/10.5281/zenodo.1211533[i])) [@cannoodt_goldstandardsinglecell_].

@@

```{r}
add_sfig(
  raw_file("overview_wrapping_v2.svg", "03-methods"),
  "overview_wrapping", 
  "A common interface for TI methods.", 
  "a) The input and output of each TI method is standardised. As input, each TI method receives either raw or normalised counts, several parameters, and a selection of prior information. After its execution, a method uses one of the seven wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. b) The specific transformations performed by each of the wrapper functions is shown in more detail.", 
  13, 
  7.5
)
```

@@

```{r}
add_stable(
  result_file("metadata.xlsx","01-datasets/01-real"), 
  "datasets", 
  "Real datasets used in this study."
)
```
## Synthetic datasets
→
To generate synthetic datasets, we used four different synthetic data simulators:

@@

Finally, to get an overall benchmarking score, we aggregated the different metrics using a geometric mean.[j][k][l][m]
←
## Method execution
Each execution of a method on a dataset was performed in a separate task as part of a gridengine job. Each task was allocated one CPU core of an Intel(R) Xeon(R) CPU E5-2665 @ 2.40GHz, and one R session was started for each task. During the execution of a method on a dataset, if the time limit (>1h) or memory limit (16GB) was exceeded, or an error was produced, a zero score was returned for that execution.
→
## Complementarity
To assess the complementarity between different methods, we first calculated for every method and dataset whether the overall score is equal or higher than 95% of the best overall score for that particular dataset. We then calculated for every method the weighted percentage of datasets which fulfilled this rule, weighted similarly as in the benchmark aggregation, and chose the best method. We iteratively added new methods until all methods were selected. For this analysis, we did not include any methods which require prior information, and only included methods which can detect at least one of the trajectory types present in the datasets.
## Scalability
To assess the scalability of each method, we started from 5 real datasets, selected using the centers from a k-medoids as discussed in the synthetic data generation section. We up- and downscaled these datasets between 10 to 100.000 cells and 10 to 100.000 features, while never going higher than 1.000.000 values in total. To generate new cells or features, we first generated a 10-nearest neighbour graph of both the cells and features from the expression space. For every new cell or feature, we used a linear combination of one to three existing cells or features, where each cell or features was given a weight sampled from $\textit{U}(0,1)$.

@@

## Stability[n][o]
We ran each method on a selection of datasets; which consisted of 25% of the datasets accounting for 15% of the total runtime, chosen such that after aggregation the overall scores still has > 0.99 correlation with the original overall ranking. We subsampled each dataset 10 times with 95% of the original cells and 95% of the original features, such that each method should produce slightly different results. 
←
## Method quality control

We created a transparent scoring scheme to check the quality of each method based on several existing tool quality and programming guidelines in literature and online (`r ref("stable", "qc_scoresheet")`). The goal of this quality control is in the first place to stimulate the improvement of current methods, and the development of user and developer friendly new methods. The quality control assessed 6 categories, each looking at several aspects, which are further divided into individual items. The availability category checks whether the method is easily available, whether the code and dependencies can be easily installed, and how the method can be used. The code quality assesses the quality of the code both from a user perspective (function naming, dummy proofing and availability of plotting functions) and a developer perspective (consistent style and code duplication). The code assurance category is frequently overlooked, and checks for code testing, continuous integration [@beaulieu-jones_reproducibility_2017] and an active support system. The documentation category checks the quality of the documentation, both externally (tutorials and function documentation) and internally (inline documentation). The behaviour category assesses the ease by which the method can be run, by looking for unexpected output files and messages, prior information and how easy the trajectory model can be extracted from the output. Finally, we also assessed certain aspects of the study in which the method was proposed, such as publication in a peer-reviewed journal, the number of datasets in which the usefulness of the method was shown, and the scope of method evaluation in the paper.

@@

[a]?
[b]we changed the title :)
[c]Try to rewrite a little bi
[d]somehing like his?
[e]TODO: rework enumeration into a paragraph
[f]something like this?
[g](Not shown), or include in Supp Fig 4?
[h]Yeah we don't include this yet, should be in sup fig 4
[i]Update when final
[j]This might give the impression of being a bit "overengineered".  Do we also show somewhere how much the resulting ranking using this aggregation would differ from a very simple "mean" ranking of the individual metrics/datasets ?
[k]No we don't do this yet. If we do this, perhaps we should just add a "benchmarking robustness" supplementary note.
[l]The results do change when we simply use a mean, but then they are very biased, eg. because 1/4 of our datasets are linear, the linear methods perform well..
[m]I adapted the text a little bit to better explain the 'why'
[n]Todo once this experiment is finished
[o]Good enough?

>>>>

---
title: "A comparison of single-cell trajectory inference methods"
output: dynbenchmark::pdf_manuscript
---

@@

`www.github.com/dynverse/dynbenchmark/tree/`r git2r::revparse_single(".","HEAD")$sha``
```{r}
n_tools <- nrow(tools)
n_tools_evaluated <- nrow(filter(tools, evaluated))
n_methods_evaluated <- nrow(filter(methods, evaluated, source %in% "tool" | id == "paga_tree"))
n_control_methods_evaluated <- nrow(filter(methods, evaluated, source %in% c("control")))
n_datasets_real <- nrow(list_datasets("real"))
n_datasets_synthetic <- nrow(list_datasets("synthetic"))
```
# Abstract

@@

→
# Results
## Trajectory inference methods
<!-- Method characterisation -->
In this study, we performed a comprehensive evaluation of  `r n_methods_evaluated` TI methods (`r ref("fig", "ti_evaluation_overview", "a")`). In order to make the outputs from different methods directly comparable to each other, we developed a common probabilistic model for representing trajectories from all possible sources (`r ref("fig", "ti_evaluation_overview", "b")`). In this model, the overall topology is represented by a network of "milestones", and the cells are placed within the space formed by each set of connected milestones. Although almost every method returned a unique set of outputs, we were able to classify these outputs into 7 distinct groups, and wrote a common output convertor for each of these groups (`r ref("fig", "results_summary", "a")` and `r ref("sfig", "overview_wrapping")`). When required, we also provided prior information to the method, the severity of which varies (`r ref("fig", "results_summary", "a")`). Namely, having to provide a start cell or an end cell is relatively easy to do and will only introduce moderate bias into the model, whereas having to provide a clustering or time course information will likely introduce heavy bias into the model.[a][b]

@@

Methods in which the topology of the model was fixed by the algorithm or by a parameter obtain a much higher stability score. As topology-free methods have more degrees of freedom, the stability of the method is also typically lower ─ with a few exceptions.[c][d]

We will discuss each evaluation criterion in more detail (`r ref("fig", "results_summary")` and `r ref("sfig", "results")`), after which we conclude with guidelines for method users and future perspectives for method developers.
```{r}
add_fig(
  fig_path = result_file("results_summary.pdf", experiment_id = "08-summary"),
  ref_id = "results_summary", 
  caption_main = glue::glue("A characterisation of the {n_methods_evaluated} methods evaluated in this study, and their overall evaluation results."), 
  caption_text = "(a) We characterised the methods according to the wrapper type, their required priors, whether the inferred topology is constrained by the algorithm (fixed) or a parameter (param), and the types of inferable topologies. The methods are grouped vertically based on the most complex trajectory type they can infer. (b) The overall results of the evaluation on four criteria: benchmarking using a reference trajectory on real and synthetic data, scalability with increasing number of cells and features, stability across dataset subsamples, and quality of the implementation.",
  width = 15,
  height = 18
)
```

```{r}
add_fig(
  fig_path = result_file("results_detailed.pdf", experiment_id = "08-summary"),
  ref_id = "results_detailed", 
  caption_main = glue::glue("Detailed results of the four main evaluation criteria: benchmarking, scalability, stability and quality control."), 
  caption_text = glue::glue("(a) The names of the methods, ordered as in {ref('fig', 'results_summary')}. (b) Results of the benchmark across metrics, dataset sources and dataset trajectory types. The performance of a method is generally more stable across dataset sources, but very variable depending on the metric and trajectory type. (c) Predicted execution times for varying numbers of cells and features (# cells × # features). Predictions were made by training a regression model after running each method on bootstrapped datasets with varying numbers of cells and features. (d) Stability results by calculating the average pairwise similarity between models inferred across multiple runs of the same method.  (e) Usability scores of the tool and corresponding manuscript, grouped per category."),
  width = 15,
  height = 18
)
```

```{r}
add_sfig(
  fig_path = result_file("results_suppfig.pdf", experiment_id = "08-summary"),
  ref_id = "results", 
  caption_main = glue::glue(""), 
  caption_text = glue::glue(""),
  width = 30,
  height = 18,
  integrate = FALSE
)
```
## Accuracy
<!-- Accuracy methods -->
We defined several possible metrics to compare a prediction to a reference trajectory (`r ref("snote", "metrics")`). Based on an analysis of their robustness and conformity to a set of rules (`r ref("snote", "metrics")`), we chose four metrics each assessing a different aspect of a trajectory (`r ref("fig", "ti_evaluation_overview", "d")`): the topology (Hamming-Ipsen-Mikhailov, `r label_metric("him")`), the quality of clustering of the assignment of cells to branches (`r label_metric("F1_branches")`), the cell positions (`r label_metric("correlation")`) and the quality of differentially expressed features  (`r label_metric("featureimp_wcor")`). The data compendium consisted of both synthetic datasets, which offer the most exact reference trajectory, and real datasets, which provide the highest biological relevance. These real datasets come from a variety of single-cell technologies, organisms, and dynamic processes, and contain several types of trajectory topologies (`r ref("stable", "datasets")`).  Some real datasets were classified as "gold standard" if the reference trajectory was not extracted from the expression data itself, such as via cellular sorting or cell mixing [@tian_scrnaseqmixologybetter_2018]. All other real datasets were classified as "silver standard". For synthetic datasets we used several data simulators, including a simulator of gene regulatory networks using a thermodynamic model of gene regulation[@schaffter_genenetweaversilicobenchmark_2011]. For each simulation, we used a real dataset as a reference, to match its dimensions, number of differentially expressed genes, drop-out rates and other statistical properties[@zappia_splattersimulationsinglecell_2017].

@@

## Scalability
While early TI methods were developed at a time where profiling more than a thousand cells was exceptional, methods now have to cope with hundreds of thousands of cells, and perhaps soon with more than ten million [@svensson_exponentialscalingsinglecell_2018]. Moreover, the recent application of TI methods on multi-omics single-cell data also showcases the increasing demands on the number of features [@cao_jointprofilingchromatin_2018]. To assess the scalability, we ran each method on up- and downscaled versions of five distinct real datasets. We then modelled the running time using a generalized additive model[@wood_thinplateregression_2003] (`r ref("sfig", "scaling", "a")`), and assessed whether the method had sublinear, linear or superlinear time complexity with respect to the number of cells and features (`r ref("sfig", "scaling", "b")`). As a control, we compared the predicted time with the actual time on all benchmarking datasets, and found that these were moderately to highly correlated (0.5-0.9) for almost every method.[e][f]

@@

## Usability
←
While not directly related to the accuracy of the inferred trajectory, it is also important to assess the quality of the implementation and how easy it is for a biological user to use [@_doesyourcode_2018]. We scored each method using a transparent checklist of important scientific and software development practices, including software packaging, documentation, automated code testing, and publication into a peer-reviewed journal (`r ref("stable", "qc_scoresheet")`).

We found that most methods fulfilled the basic criteria, such as free availability and basic code quality criteria (`r ref("fig", "results_detailed", "d")` and `r ref("sfig", "qc_overview")`). While recent methods had a slightly better quality score than older methods, several quality aspects were consistently lacking for the majority of the methods (`r ref("sfig", "qc_overview", " right")`) and we believe that these should receive extra attention from developers. Although these outstanding issues cover all five categories, code assurance and documentation in particular are problematic areas, notwithstanding several studies pinpointing these as good practices [@wilson_best_2014; @artaza_top10metrics_2016].

```{r}
add_sfig(
  fig_path = result_file("qc_overview.rds", experiment_id = "03-methods/02-tool_qc"),
  ref_id = "qc_overview", 
  caption_main = "Usability of trajectory inference methods.", 
  caption_text = glue::glue("Shown is the score given for each method on every item from our usability score sheet ({ref('stable', 'qc_scoresheet')}). Each aspect of the quality control was part of a category, and each category was weighted so that it contributed equally to the final quality score. Within each category, each aspect also received a weight depending on how often it was mentioned in a set of papers discussing good practices in tool development and evaluation. This is represented in the plot as the height on the y-axis. Top: Average QC score for each method. Right: The average score of each quality control item. Shown into more detail are those items which had an average score lower than 0.5."),
   width = 15,
   height = 18
)
```

```{r}
add_stable(
 result_file("qc_scoresheet.xlsx", "03-methods/02-tool_qc"),
  "qc_scoresheet",
paste0("→Scoring sheet for assessing usability of trajectory inference methods. Each quality aspect was given a weight based on how many times it was mentioned in a set of articles discussing best practices for tool development.←"),
  ""
)
```

@@

________________
# Online Methods
## Data and code availability
→ 
The processed real and synthetic datasets used in this study are deposited on Zenodo ([doi.org/10.5281/zenodo.1211533](https://doi.org/10.5281/zenodo.1211533[g])) [@cannoodt_goldstandardsinglecell_].

@@

```{r}
add_sfig(
  raw_file("overview_wrapping_v2.svg", "03-methods"),
  "overview_wrapping", 
  "A common interface for TI methods.", 
  "a) The input and output of each TI method is standardised. As input, each TI method receives either raw or normalised counts, several parameters, and a selection of prior information. After its execution, a method uses one of the seven wrapper functions to transform its output to the common trajectory model. This common model then allows to perform common analysis functions on trajectory models produced by any TI method. b) Illustrations of the specific transformations performed by each of the wrapper functions.", 
  13, 
  7.5
)
```

@@

```{r}
add_stable(
  result_file("real_datasets.xlsx","01-datasets/01-real"), 
  "datasets", 
  "Real datasets used in this study."
)
```
## Synthetic datasets
→
To generate synthetic datasets, we used four different synthetic data simulators:

@@

Finally, to get an overall benchmarking score, we aggregated the different metrics using a geometric mean.[h][i][j][k]
←
## Method execution
Each execution of a method on a dataset was performed in a separate task as part of a gridengine job. Each task was allocated one CPU core of an Intel(R) Xeon(R) CPU E5-2665 @ 2.40GHz, and one R session was started for each task. During the execution of a method on a dataset, if the time limit (>1h) or memory limit (16GB) was exceeded, or an error was produced, a zero score was returned for that execution.
→
## Complementarity
To assess the complementarity between different methods, we first calculated for every method and dataset whether the overall score is equal or higher than 95% of the best overall score for that particular dataset. We then calculated for every method the weighted percentage of datasets which fulfilled this rule, weighted similarly as in the benchmark aggregation, and chose the best method. We iteratively added new methods until all methods were selected. For this analysis, we did not include any methods which require prior information, and only included methods which can detect at least one of the trajectory types present in the datasets.
## Scalability
To assess the scalability of each method, we started from 5 real datasets, selected using the centers from a k-medoids as discussed in the synthetic data generation section. We up- and downscaled these datasets between 10 to 100.000 cells and 10 to 100.000 features, while never going higher than 1.000.000 values in total. To generate new cells or features, we first generated a 10-nearest neighbour graph of both the cells and features from the expression space. For every new cell or feature, we used a linear combination of one to three existing cells or features, where each cell or features was given a weight sampled from $\textit{U}(0,1)$.

@@

## Stability[l][m]
We ran each method on a selection of datasets; which consisted of 25% of the datasets accounting for 15% of the total runtime, chosen such that after aggregation the overall scores still has > 0.99 correlation with the original overall ranking. We subsampled each dataset 10 times with 95% of the original cells and 95% of the original features, such that each method should produce slightly different results. 
←
## Usability

We created a transparent scoring scheme to quantify the quality of each method based on several existing tool quality and programming guidelines in literature and online (`r ref("stable", "qc_scoresheet")`). The goal of this quality control is in the first place to stimulate the improvement of current methods, and the development of user and developer friendly new methods. The quality control assessed 6 categories, each looking at several aspects, which are further divided into individual items. The availability category checks whether the method is easily available, whether the code and dependencies can be easily installed, and how the method can be used. The code quality assesses the quality of the code both from a user perspective (function naming, dummy proofing and availability of plotting functions) and a developer perspective (consistent style and code duplication). The code assurance category is frequently overlooked, and checks for code testing, continuous integration [@beaulieu-jones_reproducibility_2017] and an active support system. The documentation category checks the quality of the documentation, both externally (tutorials and function documentation) and internally (inline documentation). The behaviour category assesses the ease by which the method can be run, by looking for unexpected output files and messages, prior information and how easy the trajectory model can be extracted from the output. Finally, we also assessed certain aspects of the study in which the method was proposed, such as publication in a peer-reviewed journal, the number of datasets in which the usefulness of the method was shown, and the scope of method evaluation in the paper.

@@

[a]Try to rewrite a little bi
[b]somehing like his?
[c]TODO: rework enumeration into a paragraph
[d]something like this?
[e](Not shown), or include in Supp Fig 4?
[f]Yeah we don't include this yet, should be in sup fig 4
[g]Update when final
[h]This might give the impression of being a bit "overengineered".  Do we also show somewhere how much the resulting ranking using this aggregation would differ from a very simple "mean" ranking of the individual metrics/datasets ?
[i]No we don't do this yet. If we do this, perhaps we should just add a "benchmarking robustness" supplementary note.
[j]The results do change when we simply use a mean, but then they are very biased, eg. because 1/4 of our datasets are linear, the linear methods perform well..
[k]I adapted the text a little bit to better explain the 'why'
[l]Todo once this experiment is finished
[m]Good enough?
